---
title: "Assignment 05"
author: "Nick Stabile"
date: "10/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(stringr)
library(lubridate)
library(sf)
library(gridExtra)
library(tidycensus)
library(httr)
library(jsonlite)
```

## Data Loading & Cleaning

```{r warning = FALSE, message = FALSE}

dc_permits_2020 <- 
  read_csv("data/Building_Permits_in_2020.csv", 
                            col_type = cols(
                              LATITUDE = col_character(),
                              LONGITUDE = col_character()
                            )) %>% 
  rename_with(tolower) %>% 
  mutate(address_id = as.character(maraddressrepositoryid))

```

## Filtering data to construction permits within six months of today

```{r warning = FALSE, message = FALSE}

dc_permits_recent_construction <- dc_permits_2020 %>% 
  filter(!is.na(latitude) & !is.na(longitude)) %>% 
  filter(permit_type_name == "CONSTRUCTION") %>% 
  filter(issue_date > today() - months(6))

table(month(dc_permits_recent_construction$issue_date))
min(dc_permits_recent_construction$issue_date)

```
##  Convert Lat/Lon to Points Geometry

```{r warning = FALSE, message = FALSE}

dc_permits_recent_construction <- st_as_sf(
  dc_permits_recent_construction,
  coords = c("longitude", "latitude"),
  crs = 4326,
  remove = FALSE)

ggplot(data = dc_permits_recent_construction) +
  geom_sf(aes(color = permit_subtype_name == "NEW BUILDING")) +
  scale_colour_manual(name = 'New Building', values = setNames(c('blue','red'),c(T, F))) +
  theme_void()

```

##  Load Census Tracts, Perform a Spatial Join, and Create Choropleth

The map shows construction permits in the last six months in DC have been highest in the central part of the District. Notably, this includes permits for anything from new buildings to additions and alterations to adding solar panels. The ratio of construction permits for new buildings was highest in a few neighborhoods east of the Anacostia River. The central parts of the city with the highest number of permits overall appear to have low numbers of new buildings. 

```{r warning = FALSE, message = FALSE}

dc_tracts <- st_read("data/Census_Tracts_in_2010.shp") %>% 
  select(TRACT, GEOID, geometry)

# Permit data doesn't have tract or geoID information, adding additional data source
dc_addresses <- st_read("data/Address_Points.csv") %>% 
  select(address_id = ADDRESS_ID, TRACT = CENSUS_TRACT)

dc_permits_recent_construction <- 
  left_join(dc_permits_recent_construction, dc_addresses, by = "address_id")

dc_permits_merged <- st_join(
  dc_permits_recent_construction, # points
  dc_tracts, # polygons
  join = st_within
)

dc_permits_merged <- st_set_geometry(dc_permits_merged, NULL)
dc_permits_merged_agg <- dc_permits_merged %>%
  group_by(GEOID) %>%
  summarize(
    num_new = sum(permit_subtype_name == "NEW BUILDING"),
    permit_count = n(),
    ratio_new = num_new/permit_count
  )

dc_permits_merged_agg <- dc_tracts %>%
  left_join(dc_permits_merged_agg, by = "GEOID")

ggplot(data = dc_permits_merged_agg) +
  geom_sf(aes(fill = permit_count)) + 
  scale_color_distiller(palette = "YlGnBu") +
  theme_void()

ggplot(data = dc_permits_merged_agg) +
  geom_sf(aes(fill = ratio_new)) + 
  scale_color_distiller(palette = "YlGnBu") +
  theme_void()

```

## Using the Census API

```{r warning = FALSE, message = FALSE}
 census_key <- read_file("data/census_api_key.txt")

# census_api_key(census_key, install = TRUE, overwrite = TRUE)

```

```{r warning = FALSE, message = FALSE}

# Exploring the ACS variables
vars <-load_variables(2017, "acs5")

# Using pivot wider for easier comparison later
dc_acs_tidy_data <- get_acs(
  geography = "tract",
  variables = c(hh_income = "B19001_001", education = "B15003_001", poverty = "B17001_001"),
  state = "DC",
  year = 2017
) %>% 
  pivot_wider(names_from = variable, values_from = c(estimate, moe))

url <- str_glue(
  "https://api.census.gov/data/2017/acs/acs5?get=NAME,B19001_001E,B15003_001E,B17001_001E&for=tract:*&in=state:11&key={census_key}"
)

# get JSON version on Census data from API
dc_acs_json <-GET(url = url, user_agent("Georgetown Univ. Student Data Collector (ncs31@georgetown.edu)."))
http_status(dc_acs_json)

# get the text of the JSON
dc_acs_json <-content(dc_acs_json, as = "text")

# convert to character matrix from JSON
dc_acs_matrix <-fromJSON(dc_acs_json)

# convert data to tibble ignoring top row which are the variable names
dc_acs_api_data <-as_tibble(dc_acs_matrix[2:nrow(dc_acs_matrix), ],.name_repair = "minimal")

# add variable names to the tibble
names(dc_acs_api_data) <- dc_acs_matrix[1, ]

# trimming columns for comparison
dc_acs_tidy_data <- dc_acs_tidy_data %>% 
  select(NAME, estimate_education, estimate_poverty, estimate_hh_income)

# Transformations for comparison
dc_acs_api_data <- dc_acs_api_data %>% 
  select(NAME, B19001_001E, B15003_001E, B17001_001E) %>% 
  rename(estimate_education = B15003_001E, 
         estimate_poverty = B17001_001E, 
         estimate_hh_income = B19001_001E) %>% 
  mutate(estimate_education = as.numeric(estimate_education),
         estimate_poverty = as.numeric(estimate_poverty),
         estimate_hh_income = as.numeric(estimate_hh_income))

all_equal(dc_acs_tidy_data, dc_acs_api_data)

```

## A Custom R Function to Query the Census API and Create a Map

```{r warning = FALSE, message = FALSE}

make_acs_map <- function(year, census_var, state_fips, county_fips) {
  
acs_map_data <- get_acs(
  geography = "tract",
  variables = census_var,
  state = state_fips,
  county = county_fips,
  year = year,
  geometry = TRUE
)

ggplot(data = acs_map_data) +
  geom_sf(aes(fill = estimate)) +
  theme_void()
  
}

# mapping median gross rent 
make_acs_map(2018, "B25064_001", 11, 1)

```

